\textls[-0]{\thetitle}

### 2023-03-05 metatheory-condensed.tex
Start working on appendices.

%In dependent type system, types are treated as terms, and should have a type.
We use Tarski-style universes~\cite{hofmann1997syntax}:
a universe~$\cU$ is not inhabited by types directly\footnote{On the contrary, 
Russel-style Universe is inhabited by types directly. 
Tarski-style universe is ``semantically more fundamental''\cite{luo2012notes} thus more friendly for meta-theoretical reasoning.},
but rather by the \emph{codes} of types:
$\codety{T}$ encodes type~$T$ (rule \ruleref{tm/c}),
and $\El{t}$ decodes~$t$ (rule \ruleref{ty/el}).
MLTT has an infinite hierarchy of universes;
we omit universe levels in the presentation for conciseness.
\rA{Why is W(t) a term and not a type?}\EDJreply{changed, please review.}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}\EDJ{changed again, please review.}

@@

The definitive version containing all the rules in \TT is stated in
terms of a meta\-language supporting quotient inductive-inductive types
(QIITs)~\cite{nordvall2010inductive,altkap2016}.
QIITs support equality constructors, which facilitate the expression of
conversion rules (e.g., \ruleref{tmeq/sub/id}).
The formalization is available as a supplemental file in Agda syntax,
though it is not checked by Agda because Agda does not support QIITs
natively.
An effort to mechanize the formalization is underway, using Agda's
\lsti{REWRITE} pragma to replace the limited uses of equality reflection.
%a \textbf{pen-and-paper} formalization
%assuming the facility of quotient inductive-inductive
%types~\cite{nordvall2010inductive,altkap2016} and extensional equality
%(equality reflection), available as a supplemental file in a
%\textbf{(pseudo-)Agda syntax}. Our formalization is directly following
%Agda syntax and type-checking (equality-checking) is done manually,
%leaving traces in the comments. We are currently working on mechanizing
%it and relying on Agda's rewriting mechanism to replace our limited
%usage of equality reflection. 
%
We also typeset a more complete set of rules than the space allows and
present it in
\ifreport \cref{sec:more-rules}.
\else a supplemental appendix.
\fi

>>>>

%In dependent type system, types are treated as terms, and should have a type.
We use Tarski-style universes~\cite{hofmann1997syntax}:
a universe~$\cU$ is not inhabited by types directly\footnote{On the contrary, 
Russel-style Universe is inhabited by types directly. 
Tarski-style universe is ``semantically more fundamental''\cite{luo2012notes} and more friendly for our meta-theoretical reasoning.},
but rather by the \emph{codes} of types:
$\codety{T}$ encodes type~$T$ (rule \ruleref{tm/c}),
and $\El{t}$ decodes~$t$ (rule \ruleref{ty/el}).
MLTT has an infinite hierarchy of universes;
we omit universe levels in the presentation for conciseness.
\rA{Why is W(t) a term and not a type?}\EDJreply{changed, please review.}\YZreply{I think the reviewer is wondering what is gained by encoding and decoding types, compared to types directly inhabiting universes.}\EDJ{changed again, please review.}

@@

The definitive version containing all the rules in \TT is stated in
terms of a meta\-language supporting quotient inductive-inductive types
(QIITs)~\cite{nordvall2010inductive,altkap2016}.
QIITs support equality constructors, which facilitate the expression of
conversion rules (e.g., \ruleref{tmeq/sub/id}).
The formalization is available as a supplemental file in Agda syntax,
though it is not checked by Agda because Agda does not support QIITs
natively.
An effort to mechanize the formalization is underway, using Agda's
\lsti{REWRITE} pragma to replace the limited uses of equality reflection.
%a \textbf{pen-and-paper} formalization
%assuming the facility of quotient inductive-inductive
%types~\cite{nordvall2010inductive,altkap2016} and extensional equality
%(equality reflection), available as a supplemental file in a
%\textbf{(pseudo-)Agda syntax}. Our formalization is directly following
%Agda syntax and type-checking (equality-checking) is done manually,
%leaving traces in the comments. We are currently working on mechanizing
%it and relying on Agda's rewriting mechanism to replace our limited
%usage of equality reflection. 
%
We also typeset a more complete set of rules than the space allows and
present it in
\ifreport \cref{sec:complete-fmltt}.
\else a supplemental appendix.
\fi
